---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "60%"
)
```

# fngr

<!-- badges: start -->
<!-- badges: end -->

The goal of `fngr` is to help maintain consistency in figures developed within ADF&G.
If you have any problems or improvements to suggest please post theme [here](https://github.com/commfish/fngr/issues).

## Installation

You can install the released version of fngr from [GitHub](https://github.com/commfish) with:

``` {r, eval = F}
# install.packages("devtools")
devtools::install_github("commfish/fngr")
```

## Example: theme_report() & theme_present()

Load the package library per usual:

```{r example, message = F}
library(fngr)
library(tidyverse)
library(scales)
```

Create data for this example

```{r}
data.frame(year = 1973:2017, age = 1:5) %>%
  mutate(y = rnorm(n(), 10, 5)) -> df
```

Base `ggplot` produces a figure with a grey background, white gridlines, and no figure border. Further the font is not in Times New Roman (standard for ADF&G reports). There are some other characteristics dealing with resolution etc that are explained in greater depth [here](http://ben-williams.github.io/updated_ggplot_figures.html) 

```{r, cache=TRUE}
ggplot(df, aes(year, age, size = y)) + 
  geom_point()  + 
  scale_size_area()
```

To make all figures have the same apperance and font set the "universal" figure theme to `theme_report()` (based off of Sean Anderson's [theme_sleek()](https://github.com/seananderson/ggsidekick)).

```{r cars}
theme_set(theme_report())
```

Create the same figure with the new theme settings. `theme_report()` is set to generate figures with a Time New Roman 11 pt font. 

```{r, cache=TRUE}
ggplot(df, aes(year, age, size = y)) + 
  geom_point()  + 
  scale_size_area()
```

R comes with only a few fonts installed.

```{r}
windowsFonts()
```

Time New Roman is not a great font for presentations, and neither is a 11 pt font. 
These can be easily adjusted. 

```{r, cache=TRUE}
theme_set(theme_report(base_family = "sans", base_size = 18))

ggplot(df, aes(year, age, size = y)) + 
  geom_point()  + 
  scale_size_area()
```

More fonts are available via the `extrafont` package. 
This was loaded on your machine along with the `fngr` package.

First you must load the available fonts (*Note that this must be done every time you update R*). 
There are a lot of fonts (likely more than necessary) and it will take a long time to load all of them. 


```{r, eval = F}
# results of font_import() not shown
extrafont::font_import()
```

Once the fonts have loaded you can explore all of the available options. 
This will generate a long list of available fonts (*Note you will need to loadfonts() during every R session that you want to use somethinging other than the base fonts*). 

```{r, message = F}
extrafont::loadfonts(device = "win")
```

Update the `base_family` font and `base_size` in the `theme_set()` if you want something different than 11 pt.
```{r, cache=TRUE}
theme_set(theme_report(base_family = "Comic Sans MS", base_size = 20))

ggplot(df, aes(year, age, size = y)) + 
  geom_point()  + 
  scale_size_area()
```

Presentation ready figure can be generated by changing the theme to `theme_present()`.
This provides an 18 pt Arial font. 
Note that axis labels may need to be adjusted to account for the increased font size. 

```{r}
xaxis = tickr(df, year, 10)

ggplot(df, aes(year, age, size = y)) + 
  geom_point() +
  scale_x_continuous(breaks = xaxis$breaks, 
                     labels = xaxis$labels) +
  theme_present()
```


Revert back to the original `theme_report()` font.

```{r}
theme_set(theme_report())
```

## Example: tickr()

The basic `ggplot` figure spaces years (x-axis) about every 5-10, depending on the timeseries length. 
Note that there are not any tick marks between the axis labels. 

```{r, cache=TRUE}
ggplot(df, aes(year, age, size = y)) + 
  geom_point()  + 
  scale_size_area()
```

One could manually define the axis breaks (ticks).

```{r, cache=TRUE}
ggplot(df, aes(year, age, size = y)) + 
  geom_point()  + 
  scale_size_area() + 
  scale_x_continuous(breaks = seq(1973, 2017, 3))
```

Or use `pretty_breaks()` from the `scales` library. 

```{r, cache=TRUE}
ggplot(df, aes(year, age, size = y)) + 
  geom_point()  + 
  scale_size_area() +
  scale_x_continuous(breaks = scales::pretty_breaks(n = 15))
```

These provide more ticks and labels, but don't place marks between the labels.

Tick marks can be adjusted manually. 

```{r, cache=TRUE}
ggplot(df, aes(year, age, size = y)) + 
  geom_point()  + 
  scale_size_area() +
  scale_x_continuous(breaks = c(1975, 1980, 2000),
                     labels = c("1975", "1980", "2000"))
```

More clearly, but with greater difficulty each year can be assigned a tickmark via `breaks =` and each tick mark is labeled via `labels =`. 
This is a viable, though cludgy method. 

```{r, cache=TRUE}
ggplot(df, aes(year, age, size = y)) + 
  geom_point()  + 
  scale_size_area() +
  scale_x_continuous(breaks = 1973:2017,
                     labels = c(rep("", 2), "1975", rep("", 4), "1980", 
                                rep("", 4), "1985", rep("", 4), "1990", 
                                rep("", 4), "1995", rep("", 4), "2000", 
                                rep("", 4), "2005", rep("", 4), "2010", 
                                rep("", 4), "2015", rep("", 2)))
```

The `tickr()` function will provide tick marks for every year (or age) and labels at a defined frequency.

Using `tickr()` the base level of labels is set at every 5th value. 

```{r, cache=TRUE}
xaxis = tickr(df, year)

ggplot(df, aes(year, age, size = y)) + 
  geom_point() +
  scale_x_continuous(breaks = xaxis$breaks, labels = xaxis$labels) 
```

This can be adjusted easily.

```{r, cache=TRUE}
xaxis = tickr(df, year, 7)

ggplot(df, aes(year, age, size = y)) + 
  geom_point() +
  scale_x_continuous(breaks = xaxis$breaks, labels = xaxis$labels) 
```

Additionally start and end values can be defined.

```{r, cache=TRUE}
xaxis = tickr(df, year, 7, start = 1973, end = 2014)

ggplot(df, aes(year, age, size = y)) + 
  geom_point() +
  scale_x_continuous(breaks = xaxis$breaks, labels = xaxis$labels) 
```

Just the start value.

```{r, cache=TRUE}
xaxis = tickr(df, year, 7, start = 1977)

ggplot(df, aes(year, age, size = y)) + 
  geom_point() +
  scale_x_continuous(breaks = xaxis$breaks, labels = xaxis$labels) 
```

Just the end value.

```{r, cache=TRUE}
xaxis = tickr(df, year, 7, end = 2010)

ggplot(df, aes(year, age, size = y)) + 
  geom_point() +
  scale_x_continuous(breaks = xaxis$breaks, labels = xaxis$labels) 
```


The start and end values can also be extended beyond the set axis by adjusting the limits.

```{r, cache=TRUE}
xaxis = tickr(df, year, 7, end = 2030)

ggplot(df, aes(year, age, size = y)) + 
  geom_point() +
  scale_x_continuous(breaks = xaxis$breaks, labels = xaxis$labels, limits = c(1970, 2030)) 
```

```{r, cache=TRUE}
xaxis = tickr(df, year, 7, start = 1970, end = 2030)

ggplot(df, aes(year, age, size = y)) + 
  geom_point() +
  scale_x_continuous(breaks = xaxis$breaks, labels = xaxis$labels, limits = c(1970, 2030)) 
```

